SDG contributions

### 2023-11-03 article/article.qmd
add figures on income groups

@@

>>>>

wb_countries_selection <- wb_countries %>% 
  select(country_code = `Country Code`, region = Region, name = `Short Name`,
         income_group = `Income Group`) %>% 
  drop_na()
```

@@

## Income regions
```{r}
papers_per_country_fos_author_pos_income_region <- papers_per_country_fos_author_pos %>%
  left_join(wb_countries_selection, by = c("country" = "country_code")) %>% 
  drop_na()

normalised_income_share_time <- papers_per_country_fos_author_pos_income_region %>% 
  group_by(SDG_label, year, income_group) %>% 
  summarise(nn = sum(n)) %>% 
  mutate(share = nn/sum(nn))

mag_2021_papers_per_country_w_income_info <- mag_2021_papers_per_country %>%
  left_join(wb_countries_selection, by = c("country" = "country_code")) %>% 
  drop_na()

mag_2021_normalised_income_share_time <- mag_2021_papers_per_country_w_income_info %>% 
  group_by(year, income_group) %>% 
  summarise(nn = sum(n)) %>% 
  mutate(share = nn/sum(nn))
```

```{r, fig.width=8.2, fig.height=4.4}
#| fig-cap: "Proportion of research from income regions on SDGs. The dashed line
#|           indicates the average across all of MAG"
#| label: fig-income_regions_over_time_with_baseline
normalised_income_share_time %>% 
  ggplot(aes(as_year(year), share, group = fix_income_group(income_group), 
             colour = fix_income_group(income_group))) +
  geom_line() +
  geom_line(data = mag_2021_normalised_income_share_time, linetype = 2) +
  facet_wrap(vars(fix_sdg(SDG_label))) +
  date_scale +
  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, y = "% of publications from world region within SDGs",
       colour = NULL)
```

### Income region specialisation
```{r}
# look at relative specialisation to identify how global south is represented in
# various SDGs
income_group_specialisation <- normalised_income_share_time %>% 
  rename(nn_sdg = nn, share_sdg = share) %>% 
  mutate(year = as.character(year)) %>% 
  left_join(mag_2021_normalised_income_share_time) %>% 
  mutate(specialisation = share_sdg / share)
```

```{r, fig.width=6, fig.height=5}
#| label: fig-specialisation-by-income-group
#| fig-cap: "Specialisation of income groups. The specialisation factor is
#| calulated by dividing the share of research towards a certain SDG by the 
#| overall share of research coming from the same income groups."
income_group_specialisation %>% 
  ggplot(aes(as_year(year), specialisation, colour = fix_sdg(SDG_label))) +
  geom_hline(yintercept = 1, colour = "grey30") +
  geom_line() +
  geom_point(size = 1.1) +
  facet_wrap(vars(fix_income_group(income_group))) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, colour = NULL, y = "Representation ratio")
```
